A meta-analysis of fungicide spray schedules to minimise yield loss from powdery mildew in Australian mungbean

### 2020-08-03 paper/paper.Rmd
Edits to intro, cross references and abbreviations`
* Copy-edit the introduction
* Fix cross-references for tables and figures
* Clean up abbreviations used, so far I've only found DSS, but there may be others that need to be added to the YAML...

#appendix: appendix.tex

@@

  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: LD 
    long: linear dichroism

@@

Mungbean was first brought to Australia in the 1930s for use as a forage or green manure crop.
However, it was only somewhat recently, the 1960s that mungbean was grown commercially [@Lawn1978; @Chauhan2018].
Prior to the 1970s the total area grown in Australia rarely exceeded 1000 hectares [@Lawn1978].
At the end of the 1980s, between 3,000 to 10,0000 hectares were being harvested for dried beans in Australia, of which mungbean is categorised by the FAO statistics data repository [@FAOSTAT].
In the decade leading up to 2018 between 35 - 86.4 thousand hectares of dried beans were planted annually.
This increase is attributed high value export markets and to the improved yields in the new cultivars [@Clarry2016].
Currently mungbean is predominantly grown in southern Queensland and northern New South Wales as a short season summer legume crop.
In 2019, approximately 90&nbsp;% of Australian mungbean was grown for export with lucrative returns of up to \$1300&nbsp;AU a tonne [@QueenslandGovernment2019].
Mungbean potentially yields up to 3 tonnes per hectare [@ThomasRobert2004], but due to high variability in yields between seasons and locations, the average Australian farm still yields less than 1 tonne per hectare [@Chauhan2018].

The high variability in yields, in part, can be attributed to a range of diseases that affect the crop [@Kelly2017a].
Two major foliar diseases, tan spot or common wilt (*Curtobacterium flaccumfaciens* pv. _flaccumfaciens_) and halo blight (*Pseudomonas savastanoi* pv. *phaseolicola*), are the main focus of the Australian Mungbean Improvement programme's resistance breeding efforts due to lack of effective chemical controls for them.
While a third disease, powdery mildew, can reduce yields by up to 40% [@Chankaew2013] in susceptible cultivars, the damage can be mitigated through an integrated disease management strategy of fungicide treatments and cultural practices.

Powdery mildew is caused by two separate genera of fungi in Australia *Podosphaera xanthii* and an unnamed *Psuedoidium sp.* (Kiss and Kelly Personal Communication).
While the pathogens are obligate parasites, secondary hosts that allow the pathogens to over-season between mungbean crops have not yet been identified. 
The disease life cycle can be as short as five days for a germinating conidia to infect the host and produce new reproductive structures, conidiophores, which produce more conidia, which disseminate to new infection sites [@Sparks2017].
Weather conditions that favour the rapid development of the disease are cool temperatures between 22&deg; and 26&deg;C, and a leaf surface that is not overly wet for infection [@Kelly2017a].

Due to the requirement for cooler temperatures for infection, planting date is an effective strategy for preventing the development of powdery mildew in mungbean [@AMAplanting].
In southern Queensland and northern New South Wales, sowing dates between late spring (November) and mid-summer (January) are advised to avoid the conducive cooler autumn temperatures, which normally commence in March.
Therefore, a delayed sowing date of late January into February increases the possibility powdery mildew will infect earlier in the crop development cycle and cause greater yield losses than in earlier sown mungbean crops.

Cultivar selection is an additional strategy to mitigate yield damage from powdery mildew. 
In Australia, quantitative resistance to powdery mildew has also been incorporated into few commercially available varieties.
Jade-AU <span class="pbr">(pbr)</span> <!-- Need PBR emblem here --> has the the best disease resistance to powdery mildew (moderately susceptible), with cv. Green Diamond also containing some notable resistance [@Sparks2017].
However commercially available mungbean cultivars in Australia lack sufficient resistance to be used as the primary powdery mildew management strategy. 
Evidence of this is shown by the 2016 field trials with the moderately susceptible cultivar, Jade-AU, where yield losses of up to 32.7 % were recorded [@SueThompson2016]. 
Quantitative disease resistance has been observed in some breeding lines overseas [@Pandey2018; @Chankaew2013], however no evidence for how they compare to Australian varietal resistance could be found.
Additional strategies to cultivar choice, such as fungicide treatments, are necessary to limit yield loss when conditions for the disease are conducive.

@@

The collection of experiments mentioned above and other unpublished trials present an excellent opportunity for a meta-analysis.
Meta-analyses are statistical tools, which can analyse a collection of experiments, that have a similar aim, and produces a more accurate estimate of the true effect being measured.
Using a meta-analysis can be useful in situations like this where several studies exist, that have the same objective but may not provide a clear answer to the question due to variation in the results within the individual studies.
The outcome of a meta-analysis provides a more accurate estimation of the true treatments effect, because it considers the amount of variance in each study and weights the influence of each studies treatment effects according to it's statistical accuracy.
Typically meta-analyses consider the effect of a single treatment against a control group across multiple independent studies [@Madden2011]. 
However meta-analyses can also consider effect differences between multiple treatments and a control group, these are called multi-variate or network meta-analyses [@MaddenEtAl2016]. 
Multi-variate meta-analyses are particularly useful when there are no direct comparisons between two treatments in any of the included studies. 
Indirect comparisons can be made between these two treatments if the both had direct comparisons with one or more treatments in common, assuming no significant bias in the studies that investigated each of the treatments and which are subject to a indirect comparison [@Jansen2011].

Considering the recent launch in 2019 of the decision support system (DSS) PowderyMildewMBM [@Diggle], which provides a cost benefit analysis to assists growers in their decision 'if' and 'when' apply fungicide; a meta-analysis to evaluate and verify the best fungicide spray schedule seemed prudent.
The aim of this meta-analysis was to determine, from a collection of unpublished studies, what spray management scenario provides the greatest yield protection from powdery mildew in Australian mungbean.

@@

```{r Table1, echo=FALSE, message=FALSE, warning=FALSE}
# Add a star to the trials which were not in the final analysis
PM_Mdat[!PM_Mdat$trial_ref %in% slimmer_PM_dat$trial_ref, "trial_ref"] <-
  sapply(PM_Mdat[!PM_Mdat$trial_ref %in% slimmer_PM_dat$trial_ref, "trial_ref"], paste0, "*")

# Build table
PM_Mdat %>%
  filter(fungicide_ai == "tebuconazole" |
           fungicide_ai == "propiconazole") %>%
  group_by(trial_ref) %>%
  summarise(
    year = unique(year),
    location = unique(location),
    planting_date = unique(planting_date),
    first_sign_disease = unique(first_sign_disease)
  ) %>%
  arrange(year) %>%
  mutate(Research_org = c(
    rep("DAF$^{a}$", times = 4),
    rep("NGA$^{b}$", times = 4),
    rep("DAF$^{a}$", times = 4),
    rep("USQ$^{c}$", times = 5)
  )) %>%
  rename(
    `Unique Trial Reference` = trial_ref,
    Year = year,
    Location = location,
    `Planting Date` = planting_date,
    `Date First Sign` = first_sign_disease,
    `Research Organisation` = Research_org
  ) %>%
  kable(format = "latex",
        booktabs = TRUE,
        escape = FALSE,
        caption="Description of fungicide application experiments.") %>%
  kable_styling() %>%
  footnote(
    alphabet = c(
      "Queensland Department of Agriculture, Fisheries and Forestry",
      "Northern Growers Alliance",
      "University of Southern Queensland"
    )
  )
```

@@

```{r Table2, echo=FALSE}
data.frame(
  Scale = 1:9,
  `Scale Description` = c(
    "No sign of powdery mildew",
    "Small colonies in lower 1/3 of the canopy, up to 75 % of the plants affected",
    "Colonies in lower 1/2 of the canopy, > 75 % of the plants affected",
    "Colonies in lower 2/3 of the canopy, up to 75 % of the plants affected",
    "Colonies in lower 2/3 of the canopy, > 75 % of the plants affected",
    "Colonies in lower 2/3 of the canopy, 100 % of the plants affected",
    "Colonies in lower 2/3 of the canopy, 100 % of plants, with some plants with colonies in the top 1/3 of the canopy",
    "Colonies to top of the plant with > 75 % of the plants affected",
    "Colonies to top of the plant with > 100 % of the plants affected and heavy leaf drop"
  )
) %>%
kable(format = "latex",
      booktabs = TRUE,
      caption="Powdery mildew severity scoring scale.") %>%
  kable_styling()
```

@@

Linear contrasts were used to produce the mean difference in effect sizes, and their respective standard errors and 95&nbsp;% confidence intervals, between each spray schedule treatment.

@@

```{r Figure1, echo = FALSE, caption = "Yield contrasts"}
YieldContrasts
```

@@

```{r Figure2, echo=FALSE, caption="Season plot"}
SeasonPlot
```

```{r Table3, echo = FALSE}
T3[, 2:5] <- round(T3[, 2:5], 4)
T3 %>%
  mutate(contrasts = gsub("-", "", contrasts)) %>%
  mutate(contrasts = gsub(" \\+", " -", contrasts)) %>%
  mutate(contrasts = gsub(" = 0", "", contrasts)) %>%
  rename(
    `Treatment contrasts` = contrasts,
    Estimate = estimates,
    `Standard Error` = se,
    `Z-score` = z_val,
    `P-value` = pval,
    ` ` = VisP
  ) %>%
  kable(format = "latex",
        booktabs = TRUE,
        caption = "Estimated grain yield contrasts between each spray schedule.") %>%
  kable_styling() %>%
  footnote(general = c("** - signficant at 0.005 level",
                       "* - signficant at 0.05 level"))
```

@@

# Data Availability
All data and code used in the preparation of this manuscript can be found in the associated research compendium.
<!-- Should we submit the raw data to a data repository like OSF and Figshare? (Paul) -->

>>>>

@@

  - short: DSS
    long: Decision support system

@@

Mungbean was first brought to Australia in the 1930s for use as a forage or green manure crop.
However, it was only somewhat recently, during the 1960s that mungbean was grown commercially [@Lawn1978; @Chauhan2018].
Prior to the 1970s the total area grown in Australia rarely exceeded 1000 hectares [@Lawn1978].
At the end of the 1980s, between 3,000 to 10,0000 hectares were being harvested for dried beans in Australia, of which mungbean is categorised by the FAO statistics data repository [@FAOSTAT].
In the decade leading up to 2018, between 35 - 86.4 thousand hectares of dried beans were planted annually.
This increase is attributed to high value export markets and to the improved yields in the new cultivars [@Clarry2016].
Currently mungbean is predominantly grown in southern Queensland and northern New South Wales as a short season summer legume crop.
In 2019, approximately 90% of Australian mungbean was grown for export with lucrative returns of up to \$1300&nbsp;AU per tonne [@QueenslandGovernment2019].
Mungbean potentially yields up to three tonnes per hectare [@ThomasRobert2004], but due to high variability in yields between seasons and locations, the average Australian farm still yields less than one tonne per hectare [@Chauhan2018].

The high variability in yields can in part be attributed to a range of diseases that affect the crop [@Kelly2017a].
Two major foliar diseases, tan spot or common wilt (*Curtobacterium flaccumfaciens* pv. _flaccumfaciens_) and halo blight (*Pseudomonas savastanoi* pv. *phaseolicola*), are the main focus of the Australian Mungbean Improvement programme's resistance breeding efforts due to lack of effective chemical controls for them.
While a third disease, powdery mildew, can reduce yields by up to 40% [@Chankaew2013] in susceptible cultivars, the damage can be mitigated through an integrated disease management strategy of fungicide treatments and cultural practices.

While the pathogens are obligate parasites, secondary hosts that allow the pathogens to over-season between mungbean crops have not yet been identified.
The disease life cycle can be as short as five days for a germinating conidia to infect the host and produce new reproductive structures, conidiophores, which produce more conidia, which disseminate to new infection sites [@Sparks2017].
Weather conditions that favour the rapid development of the disease are cool temperatures between 22&deg; and 26&nbsp;&deg;C, and a leaf surface that is not overly wet for infection [@Kelly2017a].

Due to the requirement for cool temperatures for infection, planting date is an effective strategy for preventing the development of powdery mildew in mungbean [@AMAplanting].
In southern Queensland and northern New South Wales, sowing dates between late spring (November) and mid-summer (January) are advised to avoid the conducive cooler autumn temperatures, which normally commence in March.
Therefore, delayed sowing dates of late January into February increase the possibility powdery mildew will infect earlier in the crop's life cycle and cause greater yield losses than in earlier sown mungbean crops.

Cultivar selection is an additional strategy to mitigate yield damage from powdery mildew. 
In Australia, quantitative resistance to powdery mildew has also been incorporated into few commercially available varieties.
Jade-AU <span class="pbr">(pbr)</span> <!-- Need PBR emblem here --> has the the best disease resistance to powdery mildew (moderately susceptible), with cv. Green Diamond also containing some notable resistance [@Sparks2017].
However, commercially available mungbean cultivars in Australia lack sufficient resistance to be used as the primary powdery mildew management strategy. 
Evidence of this is shown by field trials conducted in 2016 with the moderately susceptible cultivar, Jade-AU, where yield losses of up to 32.7% were recorded [@SueThompson2016].
Quantitative disease resistance has been observed in some breeding lines overseas [@Pandey2018; @Chankaew2013], however no evidence for how they compare to Australian varietal resistance could be found.
Additional strategies to cultivar choice, such as fungicide treatments, are necessary to limit yield loss when conditions for the disease are conducive.

@@

The collection of experiments mentioned above and other unpublished trials present an excellent opportunity for a meta-analysis.
Meta-analyses are statistical tools, which can analyse a collection of experiments, that have a similar aim, and produces a more accurate estimate of the true effect being measured.
Using a meta-analysis can be useful in situations like this where several studies exist that have the same objective but may not provide a clear answer to the question due to variation in the results within the individual studies.
The outcome of a meta-analysis provides a more accurate estimation of the true treatments effect because it considers the amount of variance in each study and weights the influence of each studies treatment effects according to its statistical accuracy.
Typically meta-analyses consider the effect of a single treatment against a control group across multiple independent studies [@Madden2011]. 
However meta-analyses can also consider effect differences between multiple treatments and a control group.
These are called multi-variate or network meta-analyses [@MaddenEtAl2016]. 
Multi-variate meta-analyses are particularly useful when there are no direct comparisons between two treatments in any of the included studies. 
Indirect comparisons can be made between these two treatments if the both had direct comparisons with one or more treatments in common, assuming no significant bias in the studies that investigated each of the treatments and which are subject to a indirect comparison [@Jansen2011].

Considering the recent launch in November 2019 of the decision support system (DSS) PowderyMildewMBM [@Diggle], which provides a cost benefit analysis to assists growers in their decision 'if' and 'when' apply fungicide; a meta-analysis to evaluate and verify the best fungicide spray schedule seemed prudent.
The aim of this meta-analysis was to determine, from a collection of unpublished studies, what spray management scenario provides the greatest yield protection from powdery mildew in Australian mungbean.

@@

```{r Table1, echo=FALSE, message=FALSE, warning=FALSE}
# Add a star to the trials which were not in the final analysis
PM_Mdat[!PM_Mdat$trial_ref%in% slimmer_PM_dat$trial_ref, "trial_ref"] <-
  sapply(PM_Mdat[!PM_Mdat$trial_ref%in% slimmer_PM_dat$trial_ref, "trial_ref"], paste0, "*")

# Build table
PM_Mdat%>%
  filter(fungicide_ai == "tebuconazole" |
           fungicide_ai == "propiconazole")%>%
  group_by(trial_ref)%>%
  summarise(
    year = unique(year),
    location = unique(location),
    planting_date = unique(planting_date),
    first_sign_disease = unique(first_sign_disease)
  )%>%
  arrange(year)%>%
  mutate(Research_org = c(
    rep("DAF$^{a}$", times = 4),
    rep("NGA$^{b}$", times = 4),
    rep("DAF$^{a}$", times = 4),
    rep("USQ$^{c}$", times = 5)
  ))%>%
  rename(
    `Unique Trial Reference` = trial_ref,
    Year = year,
    Location = location,
    `Planting Date` = planting_date,
    `Date First Sign` = first_sign_disease,
    `Research Organisation` = Research_org
  )%>%
  kable(format = "latex",
        booktabs = TRUE,
        escape = FALSE,
        caption="Description of fungicide application experiments.")%>%
  kable_styling()%>%
  footnote(
    alphabet = c(
      "Queensland Department of Agriculture, Fisheries and Forestry",
      "Northern Growers Alliance",
      "University of Southern Queensland"
    )
  )
```

@@

```{r Table2, echo=FALSE}
data.frame(
  Scale = 1:9,
  `Scale Description` = c(
    "No sign of powdery mildew",
    "Small colonies in lower 1/3 of the canopy, up to 75% of the plants affected",
    "Colonies in lower 1/2 of the canopy, > 75% of the plants affected",
    "Colonies in lower 2/3 of the canopy, up to 75% of the plants affected",
    "Colonies in lower 2/3 of the canopy, > 75% of the plants affected",
    "Colonies in lower 2/3 of the canopy, 100% of the plants affected",
    "Colonies in lower 2/3 of the canopy, 100% of plants, with some plants with colonies in the top 1/3 of the canopy",
    "Colonies to top of the plant with > 75% of the plants affected",
    "Colonies to top of the plant with > 100% of the plants affected and heavy leaf drop"
  )
)%>%
kable(format = "latex",
      booktabs = TRUE,
      caption="Powdery mildew severity scoring scale.")%>%
  kable_styling()
```

@@

Linear contrasts were used to produce the mean difference in effect sizes, and their respective standard errors and 95% confidence intervals, between each spray schedule treatment.

@@

```{r Figure1, echo = FALSE, fig.cap = "Yield contrasts"}
YieldContrasts
```

@@

```{r Figure2, echo=FALSE, fig.cap="Season plot"}
SeasonPlot
```

```{r Table3, echo = FALSE}
T3[, 2:5] <- round(T3[, 2:5], 4)
T3%>%
  mutate(contrasts = gsub("-", "", contrasts))%>%
  mutate(contrasts = gsub(" \\+", " -", contrasts))%>%
  mutate(contrasts = gsub(" = 0", "", contrasts))%>%
  rename(
    `Treatment contrasts` = contrasts,
    Estimate = estimates,
    `Standard Error` = se,
    `Z-score` = z_val,
    `P-value` = pval,
    ` ` = VisP
  )%>%
  kable(format = "latex",
        booktabs = TRUE,
        caption = "Estimated grain yield contrasts between each spray schedule.")%>%
  kable_styling()%>%
  footnote(general = c("** - signficant at 0.005 level",
                       "* - signficant at 0.05 level"))
```

@@

# Data Availability
All data and code used in the preparation of this manuscript can be found in the associated research compendium.
<!-- Should we submit the raw data to a data repository like OSF and Figshare? (Paul) -->
<!-- AHS - Yes, We should submit the data (as we are allowed to) to something like this, yes. See the repository I set up for the chickpea asco study, <https://zenodo.org/search?page=1&size=20&q=conceptrecid:3810826&all_versions&sort=-version> -->
